---
title: "R Assignment Notebook"
output: html_notebook
---

# Assignment problems


Make vector (7 8 6 5 4 3 2 1) 
```{r}
a = 7:1
```
Make vector ( 2 4 6 8 10 12 14 )
```{r}
b = seq(from = 2, to = 14, by = 2)
```
Make vector ( 1 2 2 3 3 3 4 4 4 4 5 5 5 5 5 )
```{r}
c = rep(c(1,2,3,4,5), c(1,2,3,4,5))
```
Make vector ( 1 2 3 1 2 3 1 2 3)
```{r}
d = rep(c(1,2,3), 3)
```
Make vector of length 100 with all entries = 3 .
```{r}
e = rep(3, 100)
```
Concatenate the first two vectors. 
```{r}
f = c(a, b)
```
%---------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------

Make a matrix ( 3 2 2; 2 3 2; 2 2 3) 
```{r}
a = matrix(c(3,2,2,2,3,2,2,2,3), nrow = 3, byrow = T)

m1 = matrix(2, nrow = 3, ncol = 3) + diag(rep(1,3))

```
Add a new row ( 2 6 8) at the end.
```{r}
b = rbind(a, c(2,6,8))
```
Add a new column ( 1 2 3 4)
```{r}
c = cbind(b, c(1,2,3,4))
```
Whats the new dimension of the matrix. 
```{r}
d = dim(c)
```
Create a matrix with no of rows = 10, ncol = 20 and all entries zero.
```{r}
e = matrix(0, nrow = 10, ncol = 20)
m2 = matrix(rep(0, 200), nrow = 10, ncol = 20)
```

%-----------------------------------------------------------------------------------------------
%-----------------------------------------------------------------------------------------------

Data structures. 

Create a data frame for the following data ( name it data_frame )

    alpha beta gamma
      1    4     7
      2    5     8
      3    6     9

```{r}
df1 = data.frame(
  alpha = c(1,2,3),
  beta = c(4,5,6),
  gamma = c(7,8,9)
)
```
Add a new column "delta" with value c(7,8,9) to the data. 
```{r}
df1$delta = c(7,8,9)

new_col = c(7,8,9)
df1 = cbind(df1, new_col)
```

Add new row ( 1 2 3 4) and ( 1 4 5 8)
```{r}
new_rows = data.frame(
  alpha = c(1,1),
  beta = c(2,3),
  gamma = c(3,5),
  delta = c(4,8)
)
df1 = rbind(df1, new_rows)
```

Create a new data frame with  column "alpha" and "gamma" ( name it new_data_frame )   
```{r}
new_data_frame = data.frame(
  alpha = df1$alpha,
  gamma = df1$gamma
)
```

Extract the data with "alpha" = 1. 
```{r}
new_data_frame[new_data_frame$alpha == 1,]
```

Create a new data frame with first two rows of new_data_frame and print it. 
```{r}
new_data_frame2 = new_data_frame[1:2,]
```


%-----------------------------------------------------------------------------------------------
%-----------------------------------------------------------------------------------------------


```{r}

library(readr)
airline = read_csv("~/Desktop/STAT 151A/STAT-151A/lab/airline.csv")

head(airline)
tail(airline)
summary(airline)

SFO_airline = airline[airline$Origin == "SFO",]
head(SFO_airline)

SFO_to_JFK_airline = SFO_airline[SFO_airline$Dest == "JFK",]
head(SFO_to_JFK_airline)
nrow(SFO_to_JFK_airline)

# 1 -> Sunday
# 7 ->Saturday
SFO_to_JFK_Sunday = SFO_to_JFK_airline[SFO_to_JFK_airline$DayOfWeek == 1,]
nrow(SFO_to_JFK_Sunday)

SFO_to_JFK_Monday = SFO_to_JFK_airline[SFO_to_JFK_airline$DayOfWeek == 2,]
nrow(SFO_to_JFK_Monday)

SFO_airline_2005 = SFO_airline[SFO_airline$Year  == 2005,]
nrow(SFO_airline_2005)

# flight from SFO to ORD in 2008 Jun
SFO_to_ORD_2008_Jun = SFO_airline[SFO_airline$Year == 2008 & SFO_airline$Dest == "ORD" & SFO_airline$Month == 6,]
nrow(SFO_to_ORD_2008_Jun)

# flight from SFO to ORD in 2005 Mar
SFO_to_ORD_2005_Mar = SFO_airline[SFO_airline$Year == 2005 & SFO_airline$Dest == "ORD" & SFO_airline$Month == 3,]
nrow(SFO_to_ORD_2005_Mar)

# input: month, year, destination
# output: airline data form SFO to destination in that month and year
get_airline_data = function(in_month, in_year, in_dest) {
  result = SFO_airline[SFO_airline$Dest == in_dest & SFO_airline$Month == in_month & SFO_airline$Year == in_year,]
  return(result)
}

nrow(get_airline_data(3, 2005, "ORD"))
```

