---
title: "Lab 1 R Assignment Notebook"
author: "Xuanpei Ouyang"
output: pdf_document
---

# Assignment problems


Make vector (7 8 6 5 4 3 2 1) 
```{r}
a = c(7:8, 6:1)
a
```

Make vector ( 2 4 6 8 10 12 14 )
```{r}
b = seq(from = 2, to = 14, by = 2)
b
```

Make vector ( 1 2 2 3 3 3 4 4 4 4 5 5 5 5 5 )
```{r}
c = rep(c(1,2,3,4,5), c(1,2,3,4,5))
c
```

Make vector ( 1 2 3 1 2 3 1 2 3)
```{r}
d = rep(c(1,2,3), 3)
d
```

Make vector of length 100 with all entries = 3 .
```{r}
e = rep(3, 100)
e
```

Concatenate the first two vectors. 
```{r}
f = c(a, b)
f
```

%---------------------------------------------------------------------------------------------

%---------------------------------------------------------------------------------------------

Make a matrix ( 3 2 2; 2 3 2; 2 2 3) 
```{r}
m1 = matrix(2, nrow = 3, ncol = 3) + diag(rep(1,3))
m1
```

Add a new row ( 2 6 8) at the end.
```{r}
m2 = rbind(m1, c(2,6,8))
m2
```

Add a new column ( 1 2 3 4)
```{r}
m3 = cbind(m2, c(1,2,3,4))
m3
```

Whats the new dimension of the matrix. 
```{r}
m3_d = dim(m3)
m3_d
```
Create a matrix with no of rows = 10, ncol = 20 and all entries zero.
```{r}
m4 = matrix(0, nrow = 10, ncol = 20)
m4
```

%-----------------------------------------------------------------------------------------------

%-----------------------------------------------------------------------------------------------

Data structures. 

Create a data frame for the following data ( name it data_frame )

    alpha beta gamma
      1    4     7
      2    5     8
      3    6     9


```{r}
data_frame = data.frame(
  alpha = c(1,2,3),
  beta = c(4,5,6),
  gamma = c(7,8,9)
)
data_frame
```

Add a new column "delta" with value c( 7,8,9) to the data. 
```{r}
new_col = data.frame(
  delta = c(7,8,9)
)
data_frame = cbind(data_frame, new_col)
data_frame
```

Add new row ( 1 2 3 4) and ( 1 4 5 8)
```{r}
new_rows = data.frame(
  alpha = c(1,1),
  beta = c(2,4),
  gamma = c(3,5),
  delta = c(4,8)
)
data_frame = rbind(data_frame, new_rows)
data_frame
```

Create a new data frame with  column "alpha" and "gamma" ( name it new_data_frame )   
```{r}
new_data_frame = data.frame(
  alpha = data_frame$alpha,
  gamma = data_frame$gamma
)
new_data_frame
```

Extract the data with "alpha" = 1. 
```{r}
new_data_frame[new_data_frame$alpha == 1,]
```

Create a new data frame with first two rows of new_data_frame and print it. 
```{r}
new_data_frame2 = new_data_frame[1:2,]
new_data_frame2
```


%-----------------------------------------------------------------------------------------------

%-----------------------------------------------------------------------------------------------


