---
title: "Examining Distributions"
author: "Seema Singh Saharan"
date: "December 22, 2016"
output: word_document
---
# Included Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(car)
library(plotrix)
library(rgl)
library(lattice)
library(ggplot2)
library(hexbin)
library(gridExtra)
library(stats)
```

# File Reading and Creation of Histogram 
```{r}
HeartAttack<-read_excel("HeartAttack.xlsx")
head(HeartAttack)
with(HeartAttack,hist(Resting_Blood_Pressure))
with(HeartAttack,hist(Resting_Blood_Pressure,breaks="FD",col="gray"))
with(HeartAttack, {
  hist(Resting_Blood_Pressure,breaks="FD",freq=FALSE,ylab="Density")
  #ylab command defines the y axis. Freq=FALSE defines Density graph instead of Frequency
  lines(density(Resting_Blood_Pressure),lwd=2)
  #lines command draws the density. lwd=2 draws double thick line.
  rug(Resting_Blood_Pressure)
  box()
  #box creates a box around the image.
})

```

# Quantile Quantile Plots
```{r}
with(HeartAttack,qqPlot(Resting_Blood_Pressure,labels = row.names(HeartAttack),id.n=3))
#The qqplot() function in the car package provides the 95% CI around the line fit to the plot.The #row name provides the Resting_Blood_Pressure being evaluated and the id.n=3 provides marking the #three most extreme points.
```

#Demonstrating normality,right skew,thick tails using QQ PLots
```{r}
qqPlot(rnorm(100,0,1))
qqPlot(rchisq(100,3))
qqPlot(rt(100,2))

```

# Boxplots 
```{r}
Boxplot(~  Resting_Blood_Pressure,data=HeartAttack)
#Parallel Box Plots
Boxplot(Resting_Blood_Pressure~Sex,data=HeartAttack, main="Blood Pressure Gender", xlab="Female(0)    Male(1)", ylab="Resting Blood Pressure")
#Boxplots of Means
mean_values<-with(HeartAttack,tapply(Resting_Blood_Pressure,Sex,mean))
standard_deviation<-with(HeartAttack,tapply(Resting_Blood_Pressure,Sex,sd))
# tapply() computes means ans sds for each gender
plotCI(1:2,mean_values,standard_deviation,xaxt="n",xlab="Gender",ylab="Blood Pressure")
# plotCI() function helps draw the graph.The parameters are coordinates on horizontal axis ,means 
# on vertical axis,standard deviations,xaxt="n" suppresses the x ticks.
lines(1:2,mean_values) 
#lines joins the means with the lines
axis(1,at=1:2,labels=names(mean_values))
# axis specifies the groups on the x axis.

```

# Scatterplots 
```{r}
scatterplot(Resting_Blood_Pressure~Age,data=HeartAttack, main="Blood Pressure Age", xlab="Age", ylab=" Resting Blood Pressure")

# scatterplot is a function in the car package that provides two smoothers on the scatterplot.
# The first smoother is the Ordinary Least Squares(OLS) which can be suppressed by reg.line=FALSE
# The second is a solid curved line which is the non parameteric regression smoother produced by # # the lowess smmoother(locally weighted scatterplot smoother) and this can be supressed by lowess # function in R.The variance is seen by the dotted line around the lowess curve.The conditional # # distribution of Resting Blood Pressure with Age can be evaluated by the vertical line between 
# dotted variance curve line.This also gives us a boxplot for both variables.

# Coded Scatter plot
scatterplot(Resting_Blood_Pressure~Age|Sex,data=HeartAttack, main="Blood Pressure Age", xlab="Age", ylab=" Resting Blood Pressure")
scatterplot(Resting_Blood_Pressure~Serum_Cholesterol,data=HeartAttack, main="Resting Blood Pressure", xlab="Serum Cholesterol", ylab="Resting Blood Pressure")

#Jittered Scatterplot
plot(jitter(Resting_Blood_Pressure,factor=2)~jitter(Age,factor=2),col="gray",cex=.5,data=HeartAttack)
with(HeartAttack, {
  abline(lm(Resting_Blood_Pressure~Age),lwd=3,lty="dashed")
  lines(lowess(Resting_Blood_Pressure,Age,f=0.2),lwd=3)
  # least square line is created by abline using the lm function
})

# Three dimensional scatterplot
scatter3d(Resting_Blood_Pressure~Age+Serum_Cholesterol,id.n=3,data=HeartAttack)
# The 3d graph shows Resting blood pressure on y axis and Age and Serum Chaolesterol on the 
# x axis and z axis.The three furthest points were identified by the Mahalanobis distances
# (Point of means) 

# Scatterplot Matrices
pairs(~ as.numeric(Resting_Blood_Pressure) + as.numeric(Serum_Cholesterol) + as.numeric(Age) ,data=HeartAttack)
```

# Hexbin 

# High Density Scatterplot with Binning

bin<-hexbin(HeartAttack$Age,HeartAttack$Resting_Blood_Pressure,xbins=50) 
plot(bin,main="Hexagonal Binning")


boxplot(Serum_Cholesterol~Sex, data = HeartAttack, lwd = 2, ylab = 'Cholesterol')
stripchart(Serum_Cholesterol~Sex, vertical = TRUE, data = HeartAttack, 
    method = "jitter", add = TRUE, pch = 20, col = 'blue')

with(HeartAttack,  
{plot(Serum_Cholesterol ~ Age)    
abline(lm(Serum_Cholesterol ~ Age), lcol="red")    
lines(lowess(Serum_Cholesterol ~ Age), lcol="blue") } ) 
 
 
#TRANSFORMATIONS using infant.mortality data set in the UN dataframe of the car package
```{r}
# Logarithm Transformations
UN <- na.omit(UN)
head(UN)
par(mfrow=c(1,2))
with(UN,plot(density(infant.mortality),xlab="Infant Mortality",main="(A)"))
with(UN,plot(density(log(infant.mortality)),
                      xlab="base e Infant Mortality",main="(B)"))
Median_Value=median(UN$infant.mortality)
Lower_Hinge=quantile(UN$infant.mortality,prob=.25)
Higher_Hinge=quantile(UN$infant.mortality,.75)
Tukey_Criterion=(Higher_Hinge-Median_Value)/(Median_Value-Lower_Hinge)
Tukey_Criterion

Infant_logfunction = log(UN$infant.mortality)
Median_Value_log=median(Infant_logfunction)
Lower_Hinge_log=quantile(Infant_logfunction,prob=.25)
Higher_Hinge_log=quantile(Infant_logfunction,prob=.75)
Tukey_Criterion_log=(Higher_Hinge_log-Median_Value_log)/(Median_Value_log-Lower_Hinge_log)
Tukey_Criterion_log
symbox(~infant.mortality,data=UN) 

#symbox uses the bc function and shows the transformation down and up the 
                      # ladder.The log function gives the best transformation

```
#Proportion Transformations using  data set in the Prestige dataframe of the car package
```{r}
Prestige <- na.omit(Prestige)
head(Prestige)
par(mfrow=c(1,3))

with(Prestige,{
  plot(density(women),main="(A) Untransformed")
  plot(density(logit(women),adjust=.75),main="(A) logit")
  plot(density(asin(sqrt(women/100)),adjust=.75),main="(A) Arcsine Square Root")

})# adjust factor smoothes the multiple modes by the decreasing the bandwidth fraction 
```
# Transformations for equalizing variance
```{r}
Ornstein <- na.omit(Ornstein)
head(Ornstein)
par(mfrow=c(1,3))

spreadLevelPlot(interlocks+1~nation,Ornstein) 
Boxplot((interlocks+1)~nation,data=Ornstein)
Boxplot(log(interlocks+1)~nation,data=Ornstein) 
# The spread is less variable across groups 
                                                  
# There are no outliers and within group spread
                                                 
 # is less skewed.
```
# Transforming Non Linearity using Tukey's Bulging Rule:
```{r}
scatterplot(infant.mortality~gdp,data=UN)  
scatterplot(log(infant.mortality)~log(gdp),data=UN)
# Notice the relationship is now linear though it might be overcorrecting this relationship.
scatterplot(sqrt(infant.mortality)~sqrt(gdp),data=UN)
# Cube root can also be used
symbox(~infant.mortality,data=UN) 
symbox(~gdp,data=UN)




