---
title: "R Assignment Notebook"
output: html_notebook
---

# Assignment problems


Make vector (7 8 6 5 4 3 2 1) 
```{r}
a = 7:1
```
Make vector ( 2 4 6 8 10 12 14 )
```{r}
b = seq(from = 2, to = 14, by = 2)
```
Make vector ( 1 2 2 3 3 3 4 4 4 4 5 5 5 5 5 )
```{r}
c = rep(c(1,2,3,4,5), c(1,2,3,4,5))
```
Make vector ( 1 2 3 1 2 3 1 2 3)
```{r}
d = rep(c(1,2,3), 3)
```
Make vector of length 100 with all entries = 3 .
```{r}
e = rep(3, 100)
```
Concatenate the first two vectors. 
```{r}
f = c(a, b)
```
%---------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------

Make a matrix ( 3 2 2; 2 3 2; 2 2 3) 

Add a new row ( 2 6 8) at the end.

Add a new column ( 1 2 3 4)

Whats the new dimension of the matrix. 

Create a matrix with no of rows = 10, ncol = 20 and all entries zero.

%-----------------------------------------------------------------------------------------------
%-----------------------------------------------------------------------------------------------

Data structures. 

Create a data frame for the following data ( name it data_frame )

    alpha beta gamma
      1    4     7
      2    5     8
      3    6     9


Add a new column "delta" with value c( 7,8,9) to the data. 

Add new row ( 1 2 3 4) and ( 1 4 5 8)

Create a new data frame with  column "alpha" and "gamma" ( name it new_data_frame )   

Extract the data with "alpha" = 1. 

Create a new data frame with first two rows of new_data_frame and print it. 



%-----------------------------------------------------------------------------------------------
%-----------------------------------------------------------------------------------------------


