---
title: "R Basics, STAT 151a"
output:
  html_notebook: default
  html_document: default
  pdf_document: default
---





Basic arithmetic operations




Addition, Subtraction, Multiplication, Division, Exponentiation



```{r}
x<- 5
y<- 2
x
y
x+y  
x*y
x-y
x/y
x^2 
```
# Vectors

R is an object oriented programming. The 3 main data structures are Vector, Matrix, Dataframe

Next we will learn how to create vectors in R. 

First method.

```{r}
x <- c(1,2,3,7,9,3) # write down each entry of vector 
x
```
The c() stands for concatenation. This method is good for creating small vectors. 

Suppose we want to create a vector which contains all the integers from 1 to 25. We can use the first method, but its time consuming to write all the entries from 1 to 25. 

Second method 

```{r}
x <- 1:25
x
```

One problem with the above method is the gap between consecutive entries is 1. So for example if we want to generate a vector (2 4 6 8 10), we cannot use the second method directly. ( One way to do it using second method is x<- 2*(1:4) ) .  Third method for creating vetors is using seq(). 


```{r}
x<- seq(from = 2, to = 12, by =2)
x
y<- seq(from=12,to=2,by=-2)
y
z<- seq(1,10,length.out = 40) # vector of length 40 with equally spaced entries, First entry is 1 alnd the last entry is 10. 
z
length(z)
```

Fourth method of generating vectors is by rep().

```{r}
x<- rep(1,25) 
x
y<- rep(1:3,3)
y
z<-rep(1:4,c(2,3,4,5))
z
```

Basic operation on a vector. Mean, Variance, Size. 

```{r}
x<-seq(from = 2, to = 12, by =0.2)
x
mean(x)
var(x)
length(x)
```

Concatenating two vectros 

```{r}
x<- 1:4
y<- 8:12
z<- c(x,y)
z
```

% ________________________Problems_______________________________________________
              
%________________________________________________________________________________




# How to create matrix in R. 

First method.  Specify a vector and tell how to fill matrix entries using the vector.

```{r}
m<- matrix(1:20,nrow=4,byrow=T) # specifies no_of_rows(nrow)  and the matrix is filled by row (byrow=T)
m
```

Fiiling entries by column

```{r}
m<- matrix(1:20,nrow=4,byrow=F) # specifies no_of_rows(nrow)  and the matrix is filled by column (byrow=F)
m
```

You can also specify the number of columns.

```{r}
m<- matrix(1:20,ncol=10,byrow=F) # specifies no_of_columns(nrow)  and the matrix is filled by column (byrow=F)
m
```

Second method is by concatenating vectors.

```{r}
a<-c(5,7,9)
b<-c(6,3,4)
c<-1:3
m<-cbind(a,b,c)
m
```

Basic matrix oprrartions.
Row-sum, Column-sum, Dimension

```{r}
m<- matrix(1:20,nrow=4,byrow=T) 
m
rowSums(m)
colSums(m)
rowMeans(m)
colMeans(m)
dim(m)
```

Addition, Subtraction, Multiplication. 

```{r}
m1<- matrix(1:25,nrow=5,byrow=T) 
m2<- matrix(3:27,nrow=5,byrow=T) 
m1
m2
m1+m2 # Has to be of same dimension
m1-m2  # Has to be of same dimension 
m1%*%m2  # note that the multiplication sign is %*% and NOT *               
```

Vector dot product and matrix elementwise product. 

```{r}
a<- 1:5
b<- 3:7

a*b # this is dot product for vectors
m1*m2 # this is matrix elementwise product 

```

Assessing vector and matrix entries.

```{r}
x<- seq(2,20,by=2) # This is shortcut for seq(from=2,to=20,by=2)
x

x[5] # 5th entry
x[1:4] # first four entry
x[c(2,3,5,8)] # 2nd 3rd 5th and 8th entry
x[-2] # all entries except 2nd one

x[-c(2,4,6)] # all entries except 2nd 4th and 6th entry


```



Getting a Sub-matrix. The method is very similar to sub-vector selestion

```{r}
m<- matrix(1:20,nrow=4,byrow=TRUE)
m

m[2,3] # the entry of m which is at the intersection of row 2 column 3. 

m[1,] # First row of m
m[,1] # First column of m
m[c(1,3),] # row 1 and 3 of matrix m
m[c(1,3),c(2,5)] # submatrix of m with row 1 row 3, column 2 , column 5

m[-2,] # all rows except row 2

```

Make a diagonal matrix

```{r}
diag(c(2,4,6))
```


%___________________PROBLEMS____________________________

%_______________________________________________________


# data frames


```{r}
n = c(2, 3, 5,7) 
s = c("aa", "bb", "cc","dd") 
b = c(TRUE, FALSE, TRUE,TRUE) 
df = data.frame(n, s, b) 
df
```


Adding a row

```{r}
new_row <- data.frame(n=9,s="ee",b=FALSE)
df<-rbind(df,new_row)
df
```

Adding a column

```{r}
new_col<- data.frame(bool=c(1,0,1,1,0))
df<-cbind(df,new_col)
df
```

Number of rows and number of columns

```{r}
nrow(df) # number of observations
ncol(df)  # number of variables 
```

Selecting few rows

```{r}
# Find the rows of df for which "bool" = 0
df_bool_0 <- df[df$bool==0,]
df_bool_0
```

Selecting few columns

```{r}
# Create a new dataset with column "n", "s" , "b"
col_names<- c("n","s","b")
new_df<- df[,col_names]
new_df
```


%_______________________________Problems________________________________

%____________________________________________________________________________

# for, while, if-else

They are similar to any other programming languages 

```{r}
for(i in 1:5){
  print(i)
}
```


```{r}
b<-2
a<- 3
if(b>a){
  print('b is greater than a')
}else{
  print("b is less than or equal to a ",quote = 0)
}
```

```{r}
a<- 5
i<- 0
while(i<a){
  cat("value of i is",i,"i is less than a \n")   # cat() can print several strings together
  i<- i+1
}
```








# Creating functions in R. 

```{r}
myfun<-function(a){  # a is the input 
  mean_a <- mean(a)
  var_a <- var(a)
  b<- (a-mean_a)/sqrt(var_a)
  return(b) # b is the output of the function
}
output<- myfun(1:30)
output
```



# Basic Plotting in R.

```{r}
head(cars)
plot(cars$speed,cars$dist,xlab="Speed",ylab = "Distance", main = "Plot of  speed vs  distance")

```
Changing plot color

```{r}
plot(cars$speed,cars$dist,xlab="Speed",ylab = "Distance", main = "Plot of  speed vs  distance" , col='red')
```

Changeing the point type and size. 'pch' controls the point type and 'cex' controls the point size ( .5 = 50% of original size ) in R plot. 


```{r}
plot(cars$speed,cars$dist,xlab="Speed",ylab = "Distance", main = "Plot of  speed vs  distance" , col='red' ,     pch = 16, cex=.5)
```

 Histogram . 

```{r}
hist(cars$speed, breaks = 8, xlab = "Car speed", main="Histogram of car speed" , col = 'orange' ) # breaks gives the number of bins ( equally spaced)
```



# Using packages


Installing packages in R. 


```{r}
#install.packages('readr')
```

Loading packages ( ot library)

```{r}
library(readr)
```


Some other data-structures in R.

List

```{r}
Student <- list(name = 'Jake', age=22, major ='Statistics' , GPA = 3.8 )
Student
```


```{r}
data<- list(x=cars[,1],y=cars[,2]) # from cars data
head(data)
```

Summary fuction

```{r}
summary(data)
```




Another function example

```{r}
my_new_fun<-function(a){
  mean_a <- mean(a)
  var_a <- var(a)
  b<- (a-mean_a)/sqrt(var_a)
  out <- list(data = a, normalizedData = b, data_mean = mean_a , data_var = var_a)
  return(out)
}
output<- my_new_fun(1:30)
output
```










