---
title: "Discussion 6"
author: "Xuanpei Ouyang (Esther)"
date: "March 21, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
```


#### (1) Do forward and Backward model selection using AIC. Are the model obtained from forward and backward selection are same?
```{r}
housing_boston = read_csv("~/Desktop/STAT 151A/STAT-151A/lab/lab6/housing_boston.csv")
colname = c("CRIM","ZN","INDUS","CHAS","NOX","RM","AGE","DIS","RAD","TAX","PTRATIO","B","LSTAT","MEDV")
names(housing_boston) = colname
summary(housing_boston)

lm_object = MEDV~CRIM+ZN+INDUS+CHAS+NOX+RM+AGE+DIS+RAD+TAX+PTRATIO+B+LSTAT
housing_boston_model = lm(lm_object, housing_boston)

AIC_forward = step(lm(MEDV ~ 1, data = housing_boston), lm_object, direction = "forward")
AIC_backward = step(housing_boston_model, MEDV ~ 1, direction = "backward")

AIC_forward
AIC_backward

AIC_forward_score = AIC(AIC_forward)
AIC_backward_score = AIC(AIC_backward)

AIC_forward_score
AIC_backward_score
```

The AIC forward model and AIC backward model are the same.



#### (2) Do forward and Backward model selection using BIC. Are the model obtained from forward and backward selection are same?
```{r}
# Find n
n = dim(housing_boston)[1]
n
BIC_forward = step(lm(MEDV ~ 1,data = housing_boston),lm_object,
                    direction = "forward",k = log(n))
BIC_backward = step(housing_boston_model, MEDV ~ 1 , direction="backward", k=log(n))

BIC_forward
BIC_backward

BIC_forward_score = BIC(BIC_forward)
BIC_backward_score = BIC(BIC_backward)

BIC_forward_score
BIC_backward_score
```

The BIC forward model and BIC backward model are different.



#### (3) Compare the forward model from Q1 and Q2.
```{r}
AIC_forward
BIC_forward
```
Therefore, we can see that the forward model from Q1 and Q2 are different.



#### (4) Compare the backward models from Q1 and Q2.
```{r}
AIC_backward
BIC_backward
```
Therefore, we can see that the backward model from Q1 and Q2 are the same.



#### (5) See if all variables obtained in the forward AIC model is significant or not ?
```{r}
AIC_forward
summary(housing_boston_model)
```
Yes, all variables obtained in the forward AIC model are significant.



#### 6.Get a 95% confidence interval for the variables for forward AIC model.
```{r}
housing_boston_model = lm(formula = MEDV~LSTAT+RM+PTRATIO+DIS+NOX+CHAS+B+ZN+
                            CRIM+RAD+TAX, data = housing_boston)
confint(housing_boston_model, c("LSTAT", "RM", "PTRATIO", "DIS", "NOX","CHAS","B","ZN","CRIM","RAD","TAX"), level = 0.95)
```
Therefore, the 95% confidence interval for the variables for forward AIC model is shown above.
