---
title: "R Notebook"
output: html_notebook
---

### Summary statistics

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(readr)
housing <- read_delim("~/Desktop/STAT 151A/STAT-151A/lab/lab6/housing.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE)

attach(housing)
```


```{r}

summary(housing)
fit = lm(price ~ sqft + bedrooms + baths)
summary(fit)
```
### Residuals, Confidence interval, fitted value
```{r}
coefficients(fit) # model coefficients
confint(fit,c("sqft","bedrooms"), level=0.95) # CIs for model parameters 
#fitted(fit) # predicted values
#residuals(fit) # residuals

```
### Getting them directly from the model 


```{r}
#fit$coefficients
#fit$residuals
summary(fit)$r.squared
summary(fit)$adj.r.squared
summary(fit)$f
```



### Adjusted R-square and Predicted R-square 

```{r, include=FALSE}
#install.packages("MPV")

```

```{r}
library("MPV")
PRESS(fit)
# OR 

pr <- residuals(fit)/(1 - lm.influence(fit)$hat)
    PRESS1 <- sum(pr^2)
    PRESS1
    
fit_anova <- anova(fit)
tss <- sum(fit_anova$"Sum Sq")
# predictive R^2
pred.r.squared <- 1 - PRESS1/(tss)
pred.r.squared
summary(fit)$adj.r.squared
```



### Dummy variable regression 

```{r}
# Dummy Variable 
fit_dummy = lm(price ~ sqft+bedrooms+baths+City)
summary(fit_dummy)
plot(fit_dummy)
```
### AIC and BIC 


```{r}
data(swiss)
attach(swiss)
lm.swiss1 <- lm(Fertility ~ Agriculture + Examination + Education + Catholic + Infant.Mortality)
lm.swiss2 <- lm(Fertility ~ Agriculture  + Education + Catholic + Infant.Mortality)
# Good model is with lowest AIC/BIC
AIC(lm.swiss1)
AIC(lm.swiss2)
# So here fit_dummy is better than fit. 
BIC(lm.swiss1)
BIC(lm.swiss2)

```



### Model selection using  AIC and BIC criteria 

```{r, message=FALSE, warning=FALSE}
attach(swiss)
library(nlme)
step(lm(Fertility ~ 1), Fertility ~ Agriculture + Examination + Education + Catholic + Infant.Mortality, direction = "forward")

n= nrow(swiss)
# Backward selection with BIC 

step(lm(Fertility ~ Agriculture + Examination + Education + Catholic + Infant.Mortality),  direction = "backward",trace=1, k=log(n))
```

### Another example 


```{r}
attach(housing)
# original model is the price 
step(lm(price ~ 1), price~ sqft + bedrooms + baths + City, direction = "forward")
# Since sqft have the lowest AIC score, we first include the sqft
# then include city
# then include b
```


```{r}
attach(housing)
n=nrow(housing)
step(lm(price~ sqft + bedrooms + baths + City),price~1 , direction = "backward",k=log(n))
# Here, if we set "k = log(n)", then the AIC shown by the function is actually the BIC value
```

#Answer the following questions using boston_housing dataset. Here MEDV is the y variable.

## 1.Do forward and Backward model selection using AIC. Are the model obtained from forward and backward selection are same?
```{r}
housing_boston <- read.csv("housing_boston.csv")
colname<-c("CRIM","ZN","INDUS","CHAS","NOX","RM","AGE","DIS","RAD","TAX",
           "PTRATIO","B","LSTAT","MEDV")
names(housing_boston)<-colname
form <- MEDV ~ CRIM +ZN+INDUS+CHAS+NOX+RM+AGE+DIS+RAD+TAX+PTRATIO+B+LSTAT
fit1 <- lm(form,housing_boston)
AICForward <- step(lm(MEDV ~1,data = housing_boston),form,direction = "forward")
AICBackward <- step(fit1,MEDV ~ 1,direction = "backward")
aicForward <- AIC(AICForward)
aicBackward <- AIC(AICBackward)
AICForward
AICBackward
```

## 2.Do forward and Backward model selection using BIC. Are the model obtained from forward and backward selection are same?
```{r}
BICForward <- step(lm(MEDV ~ 1,data = housing_boston),form,
                    direction = "forward",k = log(13))
BICBackward <- step(fit1, MEDV ~ 1 , direction = "backward",k=log(13))
bicForward <- BIC(BICForward)
bicBackward <- BIC(BICBackward)
BICForward
BICBackward

```
## 3.Compare the forward model from Q1 and Q2.
```{r}
AICForward
BICForward

aicForward
bicForward
```
## 4.Compare the backward models from Q1 and Q2.
```{r}
AICBackward
BICBackward

aicBackward
bicBackward
```
## 5.See if all variables obtained in the forward AIC model is significant or not ?
```{r}
AICForward
summary(fit1)
```
## 6.Get a 95% confidence interval for the variables for forward AIC model.
```{r}
fit <- lm(formula = MEDV ~ LSTAT + RM + PTRATIO + DIS + NOX + CHAS + 
    B + ZN + CRIM + RAD + TAX, data = housing_boston)
confint(fit,c("LSTAT", "RM", "PTRATIO", "DIS", "NOX", "CHAS","B","ZN","CRIM",
"RAD","TAX"), level=0.95)

```


```{r}
colname<-c("CRIM","ZN","INDUS","CHAS","NOX","RM","AGE","DIS","RAD","TAX","PTRATIO","B","LSTAT","MEDV")
housing_boston <- read_csv("~/Desktop/STAT 151A/STAT-151A/lab/lab6/housing_boston.csv")
names(housing_boston)<-colname

summary(housing_boston)
```

















